Informative Oceanographic Sampling using Excursion Probabilities for Multivariate Random Fields

### 2019-05-30 main_new.tex
updates

The GP mean and covariance in Eq. (\ref{prior}) are updated when more information gets available. 
In our setting, the autonomous vehicle will collect a batch of temperature and salinity measurements along a survey path. After each batch, the data are assimilated to get an updated model.
We here describe one round of updating. In Section \ref{sec:heuristics} we discuss sequential data gathering and model updating.

Denote the AUV data at $m$ locations by $\by=(y_{a,1},y_{b,1},\ldots,y_{a,m},y_{b,m})^t$, representing the temperature and salinity measurements gathered along a survey line.
Restricting attention to the grid representation, the likelihood model is described by 
\begin{equation}\label{likelihood}
\pi(\by | \bxi)=N(\bA \bxi,\bR)= \frac{1}{(2\pi)^{m}|\bR|^{\frac{1}{2}}}e^{-\frac{1}{2}(\by -\bA\bxi)^T\bR^{-1}(\by-\bA\bxi)},
\end{equation}
where $\bA$ is an $2m \times 2n$ matrix holding $1$ entries at the measurement indexes on the grid, and otherwise $0$ entries. The size $2m \times 2m$ covariance matrix $\bR$ is diagonal with entries $r_a$ and $r_b$ which are the measurement error variances of temperature and salinity observations.

The updated distribution for temperature and salinity variables on the spatial grid, given survey data $\by$, is Gaussian with mean and covariance
\begin{eqnarray}\label{gp_upd}
\bm &=& \bmu+\bSigma \bA^t (\bA \Sigma \bA^t+\bR)^{-1}(\by-\bA \bmu),  \\
\bS &=& \bSigma - \bSigma \bA^t (\bA \bSigma \bA^t+\bR)^{-1} \bA \bSigma.\nonumber
\end{eqnarray}
The updated bivariate distribution at a grid location $\bx$ is then 
\begin{equation}\label{gp_hat}
\pi 
\left(
\begin{bmatrix}
\xi_a(\bx) \\
\xi_b(\bx)
\end{bmatrix}
 |\by
 \right)
 = N \left( 
\begin{bmatrix} m_{a}(\bx)\\
m_{b}(\bx)
\end{bmatrix},\begin{bmatrix}
s_{a}^2(\bx) & s_{a,b}(\bx)  \\
s_{a,b}(\bx)  & s_{b}^2(\bx)  
\end{bmatrix}
\right),
\end{equation}
with entries extracted from the conditional mean and covariance expressions in Eq. (\ref{gp_upd}).

The conditional mean $\bm$ will be important in the following derivation in Section \ref{sec:sur}. This mean is a linear function of the data $\by$, and since the marginal distribution of the data is $\pi (\by)= N( \bA \bmu,\bA \bSigma \bA^t+\bR)$, we get
\begin{equation}\label{distxi}
\pi(\bm)= N(\bmu , \bPsi), \hspace{3mm} \bPsi=\bSigma \bA^t (\bA \bSigma \bA^t+\bR)^{-1} \bA \bSigma.  
\end{equation}
When considering only grid location $\bx$, we can extract elements from the mean vector and covariance matrix to get the bivariate distribution as follows 
\begin{equation}\label{dist_mxi}
\pi \left( \bm(\bx) \right) 
= N 
\left( 
\begin{bmatrix}
\mu_{a}(\bx) \\
\mu_{b}(\bx)
\end{bmatrix},
\begin{bmatrix}
\psi^2_{a}(\bx) & \psi_{a,b}(\bx)\\
\psi_{a,b}(\bx) & \psi^2_{b}(\bx)
\end{bmatrix}
\right).
\end{equation}

>>>>

The GP mean and covariance in Eq. (\ref{prior} \kcomment{citation
  incorrect}) are updated when more information is available. In our
setting, the autonomous vehicle will collect a batch \kcomment{what is
  a 'batch' be explicit or use another term; 'iteration'?} of
temperature and salinity measurements along a survey path; after each
batch, the data is assimilated to get an updated model. We describe
one round of updating \kcomment{where? In this section?}. In Section
\ref{sec:heuristics} we discuss sequential data gathering and the
model update process.

We denote the AUV data at $m$ locations by
$\by=(y_{a,1},y_{b,1},\ldots,y_{a,m},y_{b,m})^t$, representing the
temperature and salinity measurements gathered along a survey
transect. In the grid representation, the likelihood model is
described by:

\begin{equation}\label{likelihood}
\pi(\by | \bxi)=N(\bA \bxi,\bR)= \frac{1}{(2\pi)^{m}|\bR|^{\frac{1}{2}}}e^{-\frac{1}{2}(\by -\bA\bxi)^T\bR^{-1}(\by-\bA\bxi)},
\end{equation}

where $\bA$ is an $2m \times 2n$ matrix holding '$1$'s on the
measurement indexes on the grid, otherwise '$0$'. The size
$2m \times 2m$ covariance matrix $\bR$ has diagonal entries $r_a$ and
$r_b$ which are measurement error variances of temperature and
salinity observations.

The updated distribution for temperature and salinity variables on the
spatial grid, given survey data $\by$, is Gaussian with mean and
covariance

\begin{eqnarray}\label{gp_upd}
\bm &=& \bmu+\bSigma \bA^t (\bA \Sigma \bA^t+\bR)^{-1}(\by-\bA \bmu),  \\
\bS &=& \bSigma - \bSigma \bA^t (\bA \bSigma \bA^t+\bR)^{-1} \bA \bSigma.\nonumber
\end{eqnarray}

The updated bivariate distribution at grid location $\bx$ is then

\begin{equation}\label{gp_hat}
\pi 
\left(
\begin{bmatrix}
\xi_a(\bx) \\
\xi_b(\bx)
\end{bmatrix}
 |\by
 \right)
 = N \left( 
\begin{bmatrix} m_{a}(\bx)\\
m_{b}(\bx)
\end{bmatrix},\begin{bmatrix}
s_{a}^2(\bx) & s_{a,b}(\bx)  \\
s_{a,b}(\bx)  & s_{b}^2(\bx)  
\end{bmatrix}
\right),
\end{equation}

with entries extracted from the conditional mean and covariance
expressions in Eq. (\ref{gp_upd}).

% The conditional mean $\bm$ will be important in the following
% derivation in Section \ref{sec:sur}.
This conditional mean is a linear function of the data $\by$, and
since the marginal distribution of the data is
$\pi (\by)= N( \bA \bmu,\bA \bSigma \bA^t+\bR)$, we get
\begin{equation}\label{distxi} \pi(\bm)= N(\bmu , \bPsi), \hspace{3mm}
  \bPsi=\bSigma \bA^t (\bA \bSigma \bA^t+\bR)^{-1} \bA \bSigma.
\end{equation} When considering only grid location $\bx$, we can
extract elements from the mean vector and covariance matrix to get the
bivariate distribution as follows \begin{equation}\label{dist_mxi} \pi
  \left( \bm(\bx) \right) = N \left( \begin{bmatrix}
      \mu_{a}(\bx) \\
      \mu_{b}(\bx) \end{bmatrix}, \begin{bmatrix}
      \psi^2_{a}(\bx) & \psi_{a,b}(\bx)\\
      \psi_{a,b}(\bx) & \psi^2_{b}(\bx) \end{bmatrix} \right).
\end{equation}
